---
title: "cma-week5"
format: html
editor: visual
---

```{r}
library("readr")
library("dplyr")
library("sf")
library("ggplot2")
```

# Creating Functions

The first step in writing a function, is picking a name and assigningÂ `<- function(){}`Â to it.

```{r}
testfun <- function(){}

testfun()

class(testfun)
```

To make the function actuallyÂ *do*Â something, we need to specifyÂ *what*Â should be done within the curly bracketsÂ `{}`. The following function always prints the same statement and accepts no input values:

```{r}
testfun <- function() {
    print("this function does nothing")
}

testfun()
```

If we want the function to accept some input values, we have to define them within the round brackets. For example, I specify a variable namedÂ `sometext`Â and can call this variable within the execution.

```{r}
testfun <- function(sometext) {
    print(sometext)
}

testfun(sometext = "this function does slightly more, but still not much")
```

Let's take a more practical example. Say we want a function that calculates our age if provided with the date of our birthday. We can useÂ `Sys.time()`Â to provide today's date andÂ `difftime()`Â to calculate the time difference between today and our birthday.

```{r}
my_age <- function(birthday, output_unit) {
    difftime(Sys.time(), birthday, units = output_unit)
}

my_age(birthday = "1997-04-23", output_unit = "days")

my_age("1997-03-01", "days")
```

If we want any of our parameters to have default value, we can assign an initial value to the parameter when declaring the variables within the round brackets.

```{r}
my_age <- function(birthday, output_unit = "days") {
    difftime(Sys.time(), birthday, units = output_unit)
}

# if not stated otherwise, our function uses the unit "days"
my_age("1997-03-01")

#we can still overwrite the units

my_age("1997-03-01", "weeks")
```

# Task 1: Create your own functions:

1.  Calculate BMI based on a person's height and weight

```{r}
bmi <- function(weight, height) {
  print(weight/(height)^2)
}

bmi(70, 1.75)
```

2.  Calculate Degrees Farenheight based on Degrees Celcius

```{r}
farenheit <- function(celsius) {
  print((celsius * 9/5)+32)
}

farenheit(25)
```

3.  A function which calculates the (Euclidean) distance between two sets of coordinates (ð‘¥1, ð‘¦1 and ð‘¥2, ð‘¦2) based on the formula below:

```{r}
euc_distance <- function(x1, y1, x2, y2) {
  distance <- sqrt((x2 - x1)^2 + (y2 - y1)^2)
  return(distance)
}

```

# Task 2: Prepare Analysis

In the next tasks we will look for "meet" patterns in our wild boar data. To simplify this, we will only use a subset of our wild boar data: The individualsÂ *Rosa*Â andÂ *Sabi*Â for the timespanÂ *01.04.2015 - 15.04.2015*. You can download the dataset hereÂ [wildschwein_BE_2056.csv](https://moodle.zhaw.ch/mod/folder/view.php?id=1192125)Â and filter it with the aforementioned criteria.

```{r}
wildschwein <- read.csv("wildschwein_BE_2056.csv", header = TRUE, sep = ",")

# Konvertieren der date-Spalte in das Datum-Format
wildschwein <- wildschwein %>%
  mutate(date = as.Date(DatetimeUTC, format = "%Y-%m-%d"))

# Filtern und Konvertieren in ein sf-Objekt
d1 <- wildschwein %>%
  filter((TierName %in% c("Rosa", "Sabi")) & 
         (date >= as.Date("2015-04-01")) & 
         (date <= as.Date("2015-04-15"))) %>%
  st_as_sf(coords = c("E", "N"), crs = 2056, remove = FALSE)

```

# Task 3: Create a Join Key

The task is therfore to round the minutes ofÂ `DatetimeUTC`Â to a multiple of 15 (00, 15, 30,45) and store the values in a new column[1](https://computationalmovementanalysis.github.io/FS24/Week5/2_tasks_and_inputs.html#fn1). You can use theÂ `lubridate`Â functionÂ `round_date()`Â for this. See the examplesÂ [here](https://lubridate.tidyverse.org/reference/round_date.html)Â to see how this goes.

```{r}
library(lubridate)

d1 <- d1 |> 
  mutate(DatetimeUTC = ymd_hms(DatetimeUTC))

d1 <- d1 |> 
  mutate(DatetimeRound = round_date(DatetimeUTC, unit = "15 minutes"))

head(d1)
```

# Task 4: Measuring distance at concurrent locations

To measure the distance between concurrent locations, we need to follow the following steps.

1.  Split the `wildschwein_filter` object into one `data.frame` per animal

```{r}
d_sabi <- d1 |> 
  filter(TierName == "Sabi")

d_rosa <- d1 |> 
  filter(TierName == "Rosa")

# Konvertieren der sf-Objekte in DataFrames
rosa_df <- st_drop_geometry(d_rosa)
sabi_df <- st_drop_geometry(d_sabi)
```

1.  Join\* these datasets by the new `Datetime` column created in the last task. The joined observations are *temporally close*.

```{r}
d_joined <- inner_join(sabi_df, rosa_df, by = "DatetimeRound", suffix = c("_Sabi", "_Rosa"))
```

1.  In the joined dataset, calculate Euclidean distances between concurrent observations and store the values in a new column

```{r}
# Distance berechnen mit der Funktion aus Task 1.3

d_joined <- d_joined |> 
  mutate(distance = euc_distance(E_Sabi, N_Sabi, E_Rosa, N_Rosa))
```

1.  Use a reasonable threshold on `distance` to determine if the animals are also *spatially close* enough to constitute a *meet* (we use 100 meters). Store this Boolean information (`TRUE`/`FALSE`) in a new column

```{r}
d_joined <- d_joined |> 
  mutate(meet = distance <= 100)
```

# Task 5: Visualize Data

Now, visualize the *meets* spatially in a way that you think reasonable. For example in the plot as shows below. To produce this plot we:

-   Used the individual dataframes from `rosa` and `sabi` (from the previous task)

-   Used the joined dataset (also from the previous task), filtered to only the meets

-   Manually changed the x and y axis limits

```{r}
# Joined Dataset so filtern, dass wir nur noch die Info zu Meet haben

meet <- d_joined |> 
  filter(meet == TRUE)

# Dann den Plot erstellen

plot_meets <- ggplot() +
  #geom_path(data = d_rosa, aes(x = E, y = N), color = "blue", alpha = 0.6) +
  geom_point(data = d_rosa, aes(x = E, y = N), color = "blue", size = 1, alpha = 0.6) +
  # Sabi's Bewegungen
  #geom_path(data = d_sabi, aes(x = E, y = N), color = "red", alpha = 0.6) +
  geom_point(data = d_sabi, aes(x = E, y = N), color = "red", size = 1, alpha = 0.6) +
  # Meets
  geom_point(data = meet, aes(x = E_Rosa, y = N_Rosa), color = "green", size = 3) +
  geom_point(data = meet, aes(x = E_Sabi, y = N_Sabi), color = "green", size = 3) +
  # Achsenlimits (manuell angepasst)
  xlim(min(wildschwein$E), max(wildschwein$E)) +
  ylim(min(wildschwein$N), max(wildschwein$N)) +
  # Titel und Achsenbeschriftungen
  ggtitle("Bewegungen und Treffen von Rosa und Sabi") +
  xlab("E") + ylab("N") +
  theme_minimal()

plot_meets
  
```

# Task 6: 

```{r}
install.packages("plotly")

library(plotly)

fig <- plot_ly() |> 
  # Rosa's Bewegungen
  add_trace(data = d_rosa, x = ~E, y = ~N, z = ~DatetimeUTC, type = 'scatter3d', mode = 'lines+markers',
            line = list(color = 'blue'), marker = list(size = 1, color = 'blue'), name = 'Rosa') %>%
  # Sabi's Bewegungen
  add_trace(data = d_sabi, x = ~E, y = ~N, z = ~DatetimeUTC, type = 'scatter3d', mode = 'lines+markers',
            line = list(color = 'red'), marker = list(size = 1, color = 'red'), name = 'Sabi') %>%
  # Meets
  add_trace(data = meet, x = ~E_Rosa, y = ~N_Rosa, z = ~DatetimeUTC_Sabi, type = 'scatter3d', mode = 'markers',
            marker = list(size = 7, color = 'green', symbol = 'circle'), name = 'Meet Points') %>%
  layout(scene = list(xaxis = list(title = 'E'),
                      yaxis = list(title = 'N'),
                      zaxis = list(title = 'Time')))

fig
```
